\Title

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Methods : 2.3

Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]} $, or the emitting field of a presynaptic neuron $\synapse_\presynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]}$. As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field. It results in a weighted event stream: 
$\event_\postsynaddr = \{(\presynaddr_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}$. Unlike the binary stream of events from the signal as input, the postsynaptic neurons receives a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank = \synapticweight_\ranksyn$ when $\presynaddr_\arank = \presynaddr_\ranksyn$. This new stream of events is by nature ordered in time as events reach the soma of post-synaptic neurons. In particular, when post-synaptic neurons are activated on their soma by this spatio-temporal motif, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner. This model is defined in this subsection in all generality and the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions. In the next subsection we describe an implementation of our model adapted for the motion detection task.

@@

From the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation. Indeed, by discretization of time, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{\Nx \times \Ny \times \Ntime}$, defined in space coordinates $x$ and $y$ and time $t$. To benefit from invariance to location observed in CNN and to simplify the implementation of the model, one assumes that kernels should be similar across different positions and define a convolution operator. % The characteristics of the kernels are shared among one class of neurons such that $\forall i,j \in [1, N_\class]^2, \; \kernel^i_\class = \kernel^j_\class$. 
Here, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented as a kernel for each class $\class$ of the supervision task by the dense matrix $\kernel_\class$ of size $(\Kx, \Ky, \Ktime)$, where $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions and  $\Ktime$ is the number of delays. To keep an analogy with the HD-SNN model, for each neuron $\postsynaddr$ of position $(x, y)$ and class $\class$, $\kernel_\class$ gives the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$. Therefore, if we make the approximation that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train as input can be formalized by a 3D convolution operation. The longest synaptic delay defines the depth of the kernel $\kernel^\postsynaddr$ and all possible delays associated to the different presynaptic addresses are represented. Using this dense representation, the processing of $A$ by the model can be written as:
\begin{equation}\label{eq:kernel_b}
\forall x, y, \forall \timev \in T, \; \;
\mathcal{C}_\class(x, y, \timev)
= \sum_{\delta_x, \delta_y, \synapticdelay} \kernel_\class(\delta_x, \delta_y, \synapticdelay) \cdot A(x - \delta_x, y - \delta_y, \timev-\synapticdelay)
\end{equation}
% The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$:
% \begin{equation}\label{eq:conv}
% \forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \; \;
% \mathcal{C}_\class(\presynaddr,\timev)
% = \sum_{\ranksyn  \in [1, N_{\synapse^\postsynaddr}]} \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
% \end{equation}
This shows that $\mathcal{C}_\class$ is a spatiotemporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (see Figure~\ref{fig:model} for an illustration). Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.

The same type of spatiotemporal filtering is used as a preprocessing stage of a pattern recognition algorithm~\cite{ghosh_spatiotemporal_2019}. And in \cite{sekikawa_constant_2018}, an efficient 3D convolutional algorithm succeeds a motion estimation task. By assuming a locally constant velocity, the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity. Here, we keep the analogy with spiking neurons, and we try to use as few heuristics as possible to observe the emergence of spiking motifs in the spatiotemporal kernels. If so, this precise spatiotemporal patterns prove to be of interest for neural computations and one can assume that biological neurons make use of this information as well. 
% \note{plus de biblio sur la motion detection/segmentation ou optique flow avec les events? (reviewer 2)}
The 3D convolution represents the linear integration of the spike train as input. The activation function of our model is a softmax function implementing a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}. In our MLR model, a probability value, for each class (i.e. each motion direction), is predicted for each event at position $x, y$ and at time $\timev$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class. The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the HD-SNN model. Using the kernels, it transforms the input raster plot into a probability with the following formula:
\begin{equation}\label{eq:mlr}
    \forall x, y, \forall \timev \in T, \forall \class \in [1, N_\class], \; \;
Pr(k=\class \; \vert \;  x, y, \timev) =
\frac {\exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class) }{Z(x, y, \timev)}
\end{equation} 
where $\bias_\class$ is the bias linked to class $\class$ and $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function. 
In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike. The spiking output of the model corresponds to an event with the highest probability class.

>>>>

Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]} $, or the emitting field of a presynaptic neuron $\synapse_\presynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]}$. As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field. It results in a weighted event stream: 
$\event_\postsynaddr = \{(\presynaddr_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}$. Unlike the binary stream of events from the signal as input, the postsynaptic neurons receives a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank = \synapticweight_\ranksyn$ when $\presynaddr_\arank = \presynaddr_\ranksyn$. This new stream of events is by nature ordered in time as events reach the soma of post-synaptic neurons. In particular, when post-synaptic neurons are activated on their soma by this spatio-temporal motif, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner. This model is defined in this subsection in all generality and the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions. In the next subsection we describe an implementation of our model adapted for the motion detection task. 

@@

From the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation. Indeed, by discretization of time, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{\Nx \times \Ny \times \Ntime \times \Npol}$, defined in space coordinates $x$ and $y$, time $t$ and polarities $p$. To benefit from invariance to location observed in CNN and to simplify the implementation of the model, one assumes that kernels should be similar across different positions and define a convolution operator. % The characteristics of the kernels are shared among one class of neurons such that $\forall i,j \in [1, N_\class]^2, \; \kernel^i_\class = \kernel^j_\class$. 
Here, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented as a kernel for each class $\class$ of the supervision task by the dense matrix $\kernel_\class$ of size $(\Npol, \Kx, \Ky, \Ktime)$, where $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions and  $\Ktime$ is the number of delays. To keep an analogy with the HD-SNN model, for each neuron $\postsynaddr$ of position $(x, y)$ and class $\class$, $\kernel_\class$ gives the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$. Therefore, if we make the approximation that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train as input can be formalized by a 3D convolution operation. The longest synaptic delay defines the depth of the kernel $\kernel^\postsynaddr$ and all possible delays associated to the different presynaptic addresses are represented. Using this dense representation, the processing of $A$ by the model can be written as:
\begin{equation}\label{eq:kernel_b}
\forall x, y, \timev, \; \;
\mathcal{C}_\class(x, y, \timev)
= \sum_{p, \delta_x, \delta_y, \synapticdelay} \kernel_\class(p, \delta_x, \delta_y, \synapticdelay) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay)
\end{equation}
% The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$:
% \begin{equation}\label{eq:conv}
% \forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \; \;
% \mathcal{C}_\class(\presynaddr,\timev)
% = \sum_{\ranksyn  \in [1, N_{\synapse^\postsynaddr}]} \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
% \end{equation}
This shows that $\mathcal{C}_\class$ is a spatiotemporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (see Figure~\ref{fig:model} for an illustration). Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.

The same type of spatiotemporal filtering is used as a preprocessing stage of a pattern recognition algorithm~\cite{ghosh_spatiotemporal_2019}. And in \cite{sekikawa_constant_2018}, an efficient 3D convolutional algorithm succeeds a motion estimation task. By assuming a locally constant velocity, the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity. Here, we keep the analogy with spiking neurons, and we try to use as few heuristics as possible to observe the emergence of spiking motifs in the spatiotemporal kernels. If so, this precise spatiotemporal patterns prove to be of interest for neural computations and one can assume that biological neurons make use of this information as well. 
% \note{plus de biblio sur la motion detection/segmentation ou optique flow avec les events? (reviewer 2)}
The 3D convolution represents the linear integration of the spike train as input. The activation function of our model is a softmax function implementing a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}. In our MLR model, a probability value, for each class (i.e. each motion direction), is predicted for each event at position $x, y$ and at time $\timev$ as a softmax function of the linear combination of the list of events. % on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class. The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the HD-SNN model. 
Using the kernels, it transforms the input raster plot into a probability with the following formula:
\begin{equation}\label{eq:mlr}
    \forall x, y, \timev, \forall \class \in [1, N_\class], \; \;
Pr(k=\class \; \vert \;  x, y, \timev) =
\frac {\exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class) }{Z(x, y, \timev)}
\end{equation} 
where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function and  $\bias_\class$ is the bias linked to class $\class$. 
In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike. The spiking output of the model corresponds to an event with the highest probability class.
